---
output: html_document
title: 'How to create Giotto object with precomputed results?'
---


## 1. The Giotto class
 
A Giotto object or gobject can be created with **createGiottoObject** or **createGiottoVisiumObject** for Visium datasets.

```{r eval=FALSE}

gobject = createGiottoObject()

gvisiumobject = createGiottoVisiumObject()

```
 
createGiottoVisiumObject expects a typical Visium 10X data folder structure or a path to the h5 file, 
but createGiottoObject is more flexible and can also accept precomputed results (e.g. PCA, UMAP, network, ...)


## 2. Giotto slots

### 2.1 Required slots

The minimum requirements to make a Giotto class object are:  
- raw_exprs: matrix with raw expression values  
- spatial_locs: spatial coordinates for the cell (or spot) centroids. If this is not provided it will actually create dummy coordinates. 

### 2.2 Additional matrix slots

- norm_expr: slot to keep matrix with normalized expression values  
    - use expression_values = 'normalized'  to access this slot in downstream functions  
- norm_scaled_expr: slot to keep matrix with normalized and scaled expression values   
    - use expression_values = 'scaled'  to access this slot in downstream functions  
- custom_expr: slot to keep matrix with custom values    
    - use expression_values = 'custom'  to access this slot in downstream functions  

### 2.3 Metadata slots  

- cell_metadata:  
- gene_metadata:  

### 2.4 spatial information slots

- spatial_network:  
- spatial_network_name:  
- spatial_grid:  
- spatial_grid_name:  

### 2.5 spatial enrichment slots  

- spatial_enrichment:  
- spatial_enrichment_name:  

### 2.6 dimension reduction slots

- dimension_reduction: accepts list of dimension reduction objects

Example to add your own pre-computed PCA results    
```{r eval=FALSE}
# step1: create matrix from PCA coordinates and name the columns (PCs): Dim.1, Dim.2, Dim.3, ...
# e.g. PCA_matrix


# step2: create a pca dimension reduction object
PCA_list <- list(
     'type' = "cells", # only PCA on cells is currently supported
     'name' = "pca",   # name of your pca, 'pca' is the default name
     'reduction_method' = "pca", # needs to be pca
     'coordinates' = as.matrix(PCA_matrix), # make sure to provide a matrix
     'misc' = NULL) # slot for miscellaneous data, e.g. list(eigenvalues = ..., loadings = ...)


# step3a: add PCA_list to createGiottoObject
gobj <- createGiottoObject(...,
                           dimension_reduction=list(PCA_list), ...)

# step3b: add multiple PCA_lists to createGiottoObject
# the name parameter should be different for each PCA list
gobj <- createGiottoObject(...,
                           dimension_reduction=list(PCA_list1, PCA_list2, ...))

# step4: for UMAP or tSNE:
# same as for PCA, except for UMAP: reduction_method = 'umap' and default name = 'umap'}
# same as for PCA, except for UMAP: reduction_method = 'umap' and default name = 'umap'}

```


### 2.7 nearest neighbor network slots

- nn_network:

### 2.8 slot for images

- images:  

### 2.9 slot for offset file

- offset_file:  

### 2.10 slot for instructions  

- instructions:



