---
output: html_document
---


# How to install and work with a Giotto environment?  

To perform all potential steps and analysis in the Giotto spatial toolbox the user needs
to have a number of python modules installed. To make this process as flexible and easy 
as possible two different strategies can be used:  

$~$

#### 1. The user can install all the necessary modules themself and then the path to their python or environment (e.g. Conda) can be provided as an instruction.

```{r eval=FALSE}
library(Giotto)
```

```{r eval=FALSE}
my_instructions = createGiottoInstructions(python_path = 'your/python/path')
my_giotto_object = createGiottoObject(raw_exprs = '...',
                                      spatial_locs = '...', 
                                      instructions = my_instructions)
```

$~$

#### 2. Alternatively, the user can just install a giotto python environment using r-miniconda
as implemented in the reticulate package. In this case the environment will be automatically detected 
and no specific python path need to be provided. This is explained in more detail below:  

  
```{r eval=FALSE}
library(Giotto)
```

- installation:  
```{r eval=FALSE}
installGiottoEnvironment()
```

- re-install the giotto enviroment:  
```{r eval=FALSE}
installGiottoEnvironment(force_environment = TRUE)
```

- re-install mini-conda and enviroment:  
```{r eval=FALSE}
installGiottoEnvironment(force_miniconda = TRUE)
```

- remove Giotto environment:  
```{r eval=FALSE}
removeGiottoEnvironment()
```






