---
output: html_document
---


# How to create a Giotto object?  

#### 1. Minimum requirements  

- expression matrix  
- spatial locations  

```{r eval=FALSE}
library(Giotto)

# 1. directly from the file paths
path_to_matrix = system.file("extdata", "seqfish_field_expr.txt", package = 'Giotto')
path_to_locations = system.file("extdata", "seqfish_field_locs.txt", package = 'Giotto')

my_giotto_object = createGiottoObject(raw_exprs = path_to_matrix,
                                      spatial_locs = path_to_locations)


# 2. use an existing matrix and data.table
expression_matrix = readExprMatrix(path_to_matrix) # fast method to read expression matrix
cell_locations = data.table::fread(path_to_locations)

my_giotto_object = createGiottoObject(raw_exprs = expression_matrix,
                                      spatial_locs = cell_locations)

``` 

$~$

#### 2. Additional features   

Previous obtained information can be provided using any of the functions parameters to:  
- Add cell or gene metadata  
- add spatial networks or grids  
- add dimensions reduction  
- add giotto images  
- add offset file  
- add instructions  

Usually specificying your own instructions can be most usefull to:  
- specify a python path  
- determine output of plots  
- automatically save plots to selected directory  

```{r eval=FALSE}
library(Giotto)

# 1. directly use a path
path_to_matrix = system.file("extdata", "seqfish_field_expr.txt", package = 'Giotto')
path_to_locations = system.file("extdata", "seqfish_field_locs.txt", package = 'Giotto')

# 2. create your own instructions
path_to_python = '/usr/bin/python3' # can be something else 
working_directory = getwd() # this will use your current working directory
my_instructions = createGiottoInstructions(python_path = path_to_python,
                                           save_dir = working_directory)

# 3. create your giotto object
my_giotto_object = createGiottoObject(raw_exprs = path_to_matrix,
                                      spatial_locs = path_to_locations,
                                      cell_metadata = my_cell_metadata,
                                      gene_metadata = my_gene_metadata,
                                      instructions = my_instructions)

# 4. check which giotto instructions are associated with your giotto object
showGiottoInstructions(my_giotto_object)
```





