---
output: html_document
---


# Marker gene detection 

```{r eval=FALSE}
library(Giotto)
data("mini_giotto_single_cell")
```
This tutorial starts from a pre-computed mini Giotto object, which has already undergone pre-processing, dimensions reduction and clustering steps.  Currently provides 3 different methods to identify marker genes:  
- using a new Gini-index method  
- using Scran  
- using Mast  

Each  method can either identify marker genes between 2 selected (groups of) clusters or for each individual cluster.  

$~$

#### 1. Gini
```{r eval=FALSE}
# between 2 groups
gini_markers = findGiniMarkers(gobject = mini_giotto_single_cell,
                               cluster_column = 'leiden_clus',
                               group_1 = 1,
                               group_2 = 2)
# for each cluster
gini_markers = findGiniMarkers_one_vs_all(gobject = mini_giotto_single_cell,
                                          cluster_column = 'leiden_clus')
```

$~$

#### 2. Scran
```{r eval=FALSE}
# between 2 groups
scran_markers = findScranMarkers(gobject = mini_giotto_single_cell,
                                 cluster_column = 'leiden_clus',
                                 group_1 = 1,
                                 group_2 = 2)
# for each cluster
scran_markers = findScranMarkers_one_vs_all(gobject = mini_giotto_single_cell,
                                            cluster_column = 'leiden_clus')
```

$~$

#### 3. Mast
```{r eval=FALSE}
# between 2 groups
mast_markers = findMastMarkers(gobject = mini_giotto_single_cell,
                                cluster_column = 'leiden_clus',
                                group_1 = 1,
                                group_2 = 2)

# for each cluster
mast_markers = findMastMarkers_one_vs_all(gobject = mini_giotto_single_cell,
                                          cluster_column = 'leiden_clus')
```

$~$

#### 4. Wrapper
A general wrapper has also been created which covers all three methods:  
see **findMarkers**  and **findMarkers_one_vs_all** and specify the method parameter.  





